High-Dimensional Regularized Discriminant Analysis

### 2016-01-01 reports/paper-hdrda.Rnw
In paper updated sim description to match code

In this section we compare the competing classifiers using the simulation
design from \cite{Guo:2007te}. This design is widely used within the
high-dimensional classification literature including the studies by
\cite{Ramey:2013ji} and \cite{Witten:2011kc}. The original design comprised two
$p$-dimensional multivariate normal populations with the common block-diagonal
covariance matrix
\begin{align}
  \bm\Sigma = \bm\Sigma^{(\rho)} \oplus \bm\Sigma^{(-\rho)} \oplus \ldots \oplus \bm\Sigma^{(\rho)},\label{eq:block-sigma}
\end{align}
where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho)} \in
\mathbb{R}_{25 \times 25}$ is
\begin{align*}
\bm\Sigma_{ij}^{(\rho)} = \{ \rho^{|i - j|} \}_{1 \le i,j \le 25}.
\end{align*}
The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
gene-expression data: within each block of pathways, genes are correlated, and
the correlation decays as a function of the distance between any two genes. The
autocorrelation $\rho$ alternates between a positive and a negative value from block
to block.

Although the design is indeed standard, the simulation configuration lacks
artifacts observed commonly in real data, such as skewness and extreme
outliers. As a result, we wished to investigate the effect of outliers on the
high-dimensional classifiers. To do this, we generalized the block-diagonal
simulation configuration by sampling from a $p$-dimensional multivariate
contaminated normal distribution.  Denoting the PDF of the $p$-dimensional
multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm \Sigma)$, we
write the PDF of the $k$th class as
\begin{align}
  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma) + \epsilon N_p(\bm x | \bm \mu_k, \omega \bm \Sigma),\label{eq:contaminated-pdf}
\end{align}
where $\epsilon \in [0, 1]$ is the probability that an observation is contaminated (i.e.,
drawn from a distribution with larger variance) and $\omega > 1$ scales the
covariance matrix $\bm \Sigma$ to increase the extremity of outliers. For
$\epsilon = 0$, we have the benchmark block-diagonal simulation design from
\cite{Guo:2007te}. As $\epsilon$ is increased, the average number of outliers
is increased. In our simulation, we considered the values of $\epsilon = 0$,
$0.05$, $0.1$, $0.15$, and $0.2$. We also set $\omega = 100$.

We generated $K=3$ populations from \eqref{eq:contaminated-pdf} using the
covariance matrix $\bm \Sigma$ given in \eqref{eq:block-sigma}. We set the mean
vector of class 1 to $\bm \mu_1 = \bm 0_p$. Next, comparable to
\cite{Guo:2007te}, the first 50 features of $\bm \mu_2$ were 1/2, while the rest
were set to 0, i.e., $\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{50},
\underbrace{0, \ldots, 0}_{p - 50})$. For simplicity, we defined $\bm \mu_3 =
-\bm \mu_2$. The three populations differed in their mean vectors in the first
50 features corresponding to the first two blocks. No difference in the means
occurred in the remaining blocks.

From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
25$ for all $k$) and 1000 test observations. After training each classifier on
the training data, we classified the test data sets and computed the proportion
of mislabeled test observations to estimate the classification error rate for
each classifier. Repeating this process 250 times, we computed the average of
the error-rate estimates for each classifier. We allowed the number of features
to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
classification accuracy as the feature dimension increased while maintaining a
small sample size. Furthermore, although \cite{Guo:2007te} consider $\rho =
0.9$, we considered $\rho = 0.1$, $0.5$, and $0.9$ to explore the effect of
autocorrelation on classification accuracy, especially given that the Tong,
Pang, and Guo classifiers assume independence among features.

>>>>

In this section we compare the competing classifiers using the simulation design
from \cite{Guo:2007te}. This design is widely used within the high-dimensional
classification literature including the studies by \cite{Ramey:2013ji} and
\cite{Witten:2011kc}. First, we consider the block-diagonal covariance
matrix from from \cite{Guo:2007te}
\begin{align}
  \bm\Sigma_k = \bm\Sigma^{(\rho_k)} \oplus \bm\Sigma^{(-\rho_k)} \oplus \ldots \oplus \bm\Sigma^{(\rho_k)},\label{eq:block-sigma}
\end{align}
where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho_k)} \in
\mathbb{R}_{100 \times 100}$ is
\begin{align*}
\bm\Sigma_{ij}^{(\rho_k)} = \{ \rho_k^{|i - j|} \}_{1 \le i,j \le 100}.
\end{align*}
The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
gene-expression data: within each block of pathways, genes are correlated, and
the correlation decays as a function of the distance between any two genes. The
autocorrelation $\rho_k$ alternates between a positive and a negative value from block
to block. The original design from \cite{Guo:2007te} comprised two $p$-dimensional
multivariate normal populations with a common block-diagonal covariance
matrix where $\bm\Sigma_1 = \bm\Sigma_2$.

Although the design is indeed standard, the simulation configuration lacks
artifacts observed commonly in real data, such as skewness and extreme
outliers. As a result, we wished to investigate the effect of outliers on the
high-dimensional classifiers. To do this, we generalized the block-diagonal
simulation configuration by sampling from a $p$-dimensional multivariate
contaminated normal distribution.  Denoting the PDF of the $p$-dimensional
multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm \Sigma)$, we write
the PDF of the $k$th class as
\begin{align}
  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma_k) + \epsilon N_p(\bm x | \bm \mu_k, \omega \bm \Sigma_k),\label{eq:contaminated-pdf}
\end{align}
where $\epsilon \in [0, 1]$ is the probability that an observation is
contaminated (i.e., drawn from a distribution with larger variance) and $\omega
> 1$ scales the covariance matrix $\bm \Sigma_k$ to increase the extremity of
outliers. For $\epsilon = 0$, we have the benchmark block-diagonal simulation
design from \cite{Guo:2007te}. As $\epsilon$ is increased, the average number of
outliers is increased. In our simulation, we considered the values of $\epsilon
= 0$, $0.05$, $\ldots$, $0.50$. We also set $\omega = 100$.

We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
\Sigma_k$ given in \eqref{eq:block-sigma}. We set the mean vector of class 1 to
$\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
features of $\bm \mu_2$ were 1/2, while the rest were set to 0, i.e., $\bm \mu_2
= (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p -
  100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
populations differed in their mean vectors in the first 100 features
corresponding to the first block. No difference in the means occurred in the
remaining blocks.

From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
25$ for all $k$) and 10,000 test observations. After training each classifier on
the training data, we classified the test data sets and computed the proportion
of mislabeled test observations to estimate the classification error rate for
each classifier. Repeating this process 500 times, we computed the average of
the error-rate estimates for each classifier. We allowed the number of features
to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
classification accuracy as the feature dimension increased while maintaining a
small sample size. Furthermore, although \cite{Guo:2007te} consider $\rho_k =
0.9$ for all $k$, we considered $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
to explore the more realistic assumption of unequal covariance matrices.
